Master's Project Journal

### 2019-03-23 conf_paper/paper.tex
new LF test module without fractions & 4 page + ref draft of paper

\begin{abstract}
    This document is a model and instructions for \LaTeX.
    This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
    or Math in Paper Title or Abstract.
\end{abstract}

@@

    FPGA, ADPLL, Simulation, Prototype

@@

An FPGA as a prototyping tool has seen use in the work of Dimitri Galayko and his research group in UPMC Sorbonne and in Elena Blokhina's team in University College Dublin, and has been used in the testing of established designs prior to their implementation in custom silicon\cite{zianbetov2013phd,shan2014phd}, the creation of realistic models for use in high level simulations\cite{theboys2019} and the exploration of new designs for digital blocks before progressing on to true digital design on a gate level.  %TODO references

@@

\subsubsection*{Design 1}
The first of these is a linear in period design using an accumulator architecture, with a variable reload value on overflow. Here the control code modifies the reload value around a bias point and thus adjusts the time taken for overflow to be reached with the oscillator's period given as $T_{osc} = (2^{\textrm{width}} - (\textrm{bias}+\textrm{code}))\times\textrm{clock period}$. Period linearity is a desirable characteristic as it mimics the behaviour of many mixed-signal implementations, and is the key use case of this method. The choice of the accumulator's Most Significant Bit (MSB) as the output signal however highlights the drawbacks of this method as less than half of the control codes are feasible. As the minimum period is set by the number of unused bits and the maximum period fixed at twice the minimum the frequency and tuning range combinations are very limited.
In addition the wave output by this oscillator is only square at the maximum period and control code changes vary the duty cycle of the signal.

\subsubsection*{Design 2}
If frequency linearity is a desired characteristic of the oscillator an accumulator based, FPGA clock driven design exists for this use case also. 
Here the control code $k$ is added to the value stored in the accumulator at each rising edge of the FPGA clock until overflow is reached at $2^n-1$ where $n$ is the bit width of the counter thus valuing each control code increment at $\frac{\textrm{FPGA clock}}{2^{\textrm{width}}}$ Hz. Once again the MSB of the accumulator is used as the output signal, however with this design the MSB will spend an equivalent duration at both logic high and low.
Again this design is better suited to lower frequencies, where the output of the DCO is orders of magnitude lower than the clock signal driving it.

\subsubsection*{Design 3}
The third and final potential DCO design is an inverter ring/chain which makes use of the delay through inverter primitives, thus bearing significant resemblance to DCOs used on ASICs. On an FPGA the lack of precise control over the layout of blocks means the exact value of the delay is uncertain and may vary noticeably from one implementation to another, however this misalignment and the asynchronous behaviour stemming from the lack of an FPGA driven clock increase the similarity between this design and an ASIC. The period is set by the number of inverters at $2\times(\textrm{num inverters})\times(\textrm{delay})$, and is varied by adding or removing a pair of inverters for every control code increment. This design works best when the desired period is within a couple orders of magnitude of the delay through each pair of inverters, beyond this point the number of look-up tables and routing resources required to implement the design becomes unsustainable.

@@

\subsubsection*{Design B}
Better suited for use with the inverter based oscillator is a system that mimics the behaviour of a tapped delay line working asynchronously to the FPGA clock. As the FPGA clock is not used, and that of the ADPLL cannot be used to measure itself an unclocked sign detection circuit is required, an example of which can be seen in Figure \ref{fig:segn_pdetdl}. This is nearly identical to sign detection methods used on an ASIC with the arbitration circuit replaced by a second SR Latch which alleviates the issue of metastability.\\
A tapped delay line can be emulated on an FPGA through the use of inverter pairs due to their short and approximately known delay placed between each tap.
The resolution then is the propagation time through the pair, although this will once again depend on the layout specifics, which was approximately 600 picoseconds on an Artix-7. This is significantly lower than the 4 nanoseconds achieved using the previous design and thus more suitable for use with an inverter ring oscillator as this resolution is better than the time resolution of the oscillator. Figure \ref{fig:num_pdetdl} illustrates the configuration of the circuit used to emulate the function of a delay line that will be used in this paper.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{../new_pdet1}
    \caption{Design B Sign Detection RTL Diagram.}
    \label{fig:segn_pdetdl}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{../new_pdet2}
    \caption{Tapped Delay Line Emulation RTL Diagram.}
    \label{fig:num_pdetdl}
\end{figure}
\section{Comparison of Designs}
This paper will examine the performance of a number of combinations of these blocks and the performance thereof, specifically oscillator 2 with phase detector A and oscillator 3 with detectors A \& B. For each test there were four ADPLLs on the FPGA arranged in a Cartesian Grid and as such each ADPLL will be referred to using matrix indexing corresponding to their grid location. For each configuration three different modes of operation were examined: Each ADPLL locking to the same 5 MHz external reference, each ADPLL connected to the left and above neighbours in a ``uni-directional'' network and finally a true ADPLL network where all neighbours are connected. For comparing each configuration gains which had previously been determined to provide the best jitter performance were chosen. The data was captured using an Agilent MSO7054 A at 4 GSa/s. Each point of data is the mean of five 200 microsecond long captures which were then analysed in Matlab, and Table \ref{table:b_c_perf} displays the results of this analysis.\\
The first columns of the table relate to the cycle-to-cycle jitter of the system, the difference between sequential distances, and the second portion gives the Time Interval Error, or difference between the actual timing of the edge and the ``ideal'' time as set by the reference. This combination of measurements will show how well each design performs the task of staying consistent from cycle to cycle both in relation to itself and the system at large.

\begin{table}[!ht]
	\begin{center}
		\begin{small}
			\begin{tabular}{cc|c|c|c|c|c|c|}           
				\cline{3-8}
				&& \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Time Interval Error. (ns)} \T\\
				\cline{3-8}	
				&&PLL 11&PLL 12&PLL 22&PLL 11&PLL 12&PLL 22\T\\
				\hline
				\multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
				\multicolumn{2}{|r|}{Free PLLs}&1.5988&1.6178&1.6220&&& \T\\
				\multicolumn{2}{|r|}{Uni-dir.}&1.9804&2.0316&2.1317&&& \T\\
				\multicolumn{2}{|r|}{Bi-dir.}&2.0124&2.0681&2.0732&&& \T\\
				\hline
				\multicolumn{2}{|l|}{Osc 3, PDET A}&-&-&-&-&-&-\T\\
				\multicolumn{2}{|r|}{Free PLLs}&0.62499 $\mu$s&0.75919 $\mu$s&0.79049  $\mu$s&&& \T\\
				\multicolumn{2}{|r|}{Uni-dir.}&0.76822 $\mu$s&0.9009 $\mu$s&1.1407 $\mu$s&&& \T\\
				\multicolumn{2}{|r|}{Bi-dir.}&0.70393 $\mu$s&0.70201 $\mu$s&1.241 $\mu$s&&& \T\\
				\hline
				\multicolumn{2}{|l|}{Osc 3, PDET B}&-&-&-&-&-&-\T\\
				\multicolumn{2}{|r|}{Free PLLs}&4.1680&3.9803&5.0741&&& \T\\
				\multicolumn{2}{|r|}{Uni-dir.}&4.3820&5.7315&6.2336 &&& \T\\
				\multicolumn{2}{|r|}{Bi-dir.}&5.6137&7.1831&7.1612 &&& \T\\
				\hline
				\B
			\end{tabular}
		\end{small}
		\caption{}
		\label{table:b_c_perf}
	\end{center}
\end{table}

From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation. In this system the FPGA clock is at 258 MHz, the maximum possible, thus giving a detection resolution of 4 nanoseconds. This is 4 times greater than the period resolution of approximately 1.2 nano seconds of the oscillator, and requires the gains of the PI filter to be increased fourfold in order to respond appropriately to the error detected. The problems of relatively inaccurate detection and increased gains combine to produce a system that is only able to achieve poor locking performance.
Both other combinations work well however, delivering a level of jitter that is both low yet allows an indication how impactful variations in architecture or control parameters to be seen visually on an oscilloscope.\\
\begin{figure}[h]
	\centering
	\includegraphics[width=0.25\textwidth]{distrib_ring}
	\caption{Design B Sign Detection RTL Diagram.}
	\label{fig:dist1}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.25\textwidth]{distrib_pa}
	\caption{Tapped Delay Line Emulation RTL Diagram.}
	\label{fig:dist2}
\end{figure}
Figures \ref{fig:dist1} \& \ref{fig:dist2} paint a picture of the unrealistic nature of the FPGA clocked designs as in comparison to the ring based design each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment. This effects the suitability of the FPGA clocked designs for the improvement of simulation models or the examination of how a novel block maybe impact the behaviour of the network.
\section*{Conclusion}
An FPGA is a suitable environment for the prototyping of ADPLL networks and can fill a number of roles depending on the designer's requirements. However the designer should bear in mind the platform's limitations especially when attempting to use a frequency that is not orders of magnitude lower than the maximum clock available from the FPGA's clock manager. Inverter based and FPGA clocked designs both have a niche in which they are the preferred option, with the clocked design better suiting replication of an existing design at scaled down frequencies and the inverter based designs better in the modelling and analysis of ADPLL network behaviour.
\bibliography{conf.bib} 
\bibliographystyle{IEEEtran}

>>>>

\begin{abstract}
    This paper introduces and compares the implementation of a number of FPGA based ADPLL network prototyping architectures. Comparison is then made between the expected performance and role of each method as a development tool.
\end{abstract}

@@

    FPGA, ADPLL, Simulation, Prototyping

@@

An FPGA as a prototyping tool has seen use in UPMC Sorbonne and University College Dublin, and has been used in the testing of established designs prior to their implementation in custom silicon\cite{zianbetov2013phd,shan2014phd}, the validation of realistic models for use in high level simulations\cite{theboys2019} and the exploration of new designs for digital blocks before progressing on to true digital design on a gate level.  %TODO references

@@

%\subsubsection*{Design 1}
%The first of these is a linear in period design using an accumulator architecture, with a variable reload value on overflow. Here the control code modifies the reload value around a bias point and thus adjusts the time taken for overflow to be reached with the oscillator's period given as $T_{osc} = (2^{\textrm{width}} - (\textrm{bias}+\textrm{code}))\times\textrm{clock period}$. Period linearity is a desirable characteristic as it mimics the behaviour of many mixed-signal implementations, and is the key use case of this method. The choice of the accumulator's Most Significant Bit (MSB) as the output signal however highlights the drawbacks of this method as less than half of the control codes are feasible. As the minimum period is set by the number of unused bits and the maximum period fixed at twice the minimum the frequency and tuning range combinations are very limited.
%In addition the wave output by this oscillator is only square at the maximum period and control code changes vary the duty cycle of the signal.

\subsubsection*{Design 1}
If frequency linearity is a desired characteristic of the oscillator an accumulator based, FPGA clock driven design exists for this use case also. 
Here the control code $k$ is added to the value stored in the accumulator at each rising edge of the FPGA clock until overflow is reached at $2^n-1$ where $n$ is the bit width of the counter thus valuing each control code increment at $\frac{\textrm{FPGA clock}}{2^{\textrm{width}}}$ Hz. Once again the MSB of the accumulator is used as the output signal, however with this design the MSB will spend an equivalent duration at both logic high and low.
Again this design is better suited to lower frequencies, where the output of the DCO is orders of magnitude lower than the clock signal driving it.

\subsubsection*{Design 2}
The third and final potential DCO design is an inverter ring/chain which makes use of the delay through inverter primitives, thus bearing significant resemblance to DCOs used on ASICs. On an FPGA the lack of precise control over the layout of blocks means the exact value of the delay is uncertain and may vary noticeably from one implementation to another, however this misalignment and the asynchronous behaviour stemming from the lack of an FPGA driven clock increase the similarity between this design and an ASIC. The period is set by the number of inverters at $2\times(\textrm{num inverters})\times(\textrm{delay})$, and is varied by adding or removing a pair of inverters for every control code increment. This design works best when the desired period is within a couple orders of magnitude of the delay through each pair of inverters, beyond this point the number of look-up tables and routing resources required to implement the design becomes unsustainable.\\
A third potential design exists, similar to Oscillator 1 in which the control code varies the reload value of the accumulator at overflow, however this design will not be discussed in this paper.

@@

\subsubsection*{Design B}
Better suited for use with the inverter based oscillator is a system that mimics the behaviour of a tapped delay line working asynchronously to the FPGA clock. As the FPGA clock is not used, and that of the ADPLL cannot be used to measure itself an unclocked sign detection circuit is required, an example of which can be seen in Figure \ref{fig:segn_pdetdl}. This is nearly identical to sign detection methods used on an ASIC with the arbitration circuit replaced by a second SR Latch which alleviates the issue of metastability.\\
A tapped delay line can be emulated on an FPGA through the use of inverter pairs due to their short and approximately known delay placed between each tap.
The resolution then is the propagation time through the pair, although this will once again depend on the layout specifics, which was approximately 600 picoseconds on an Artix-7. This is significantly lower than the 4 nanoseconds achieved using the previous design and thus more suitable for use with an inverter ring oscillator as this resolution is better than the time resolution of the oscillator. Figure \ref{fig:num_pdetdl} illustrates the configuration of the circuit used to emulate the function of a delay line that will be used in this paper.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{../new_pdet1}
    \caption{Design B Sign Detection RTL Diagram.}
    \label{fig:segn_pdetdl}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{../new_pdet2}
    \caption{Tapped Delay Line Emulation RTL Diagram.}
    \label{fig:num_pdetdl}
\end{figure}
\section{Comparison of Designs}
This paper will examine the performance of a number of combinations of these blocks and the performance thereof, specifically oscillator 1 with phase detector A and oscillator 2 with detectors A \& B. For each test there were four ADPLLs on the FPGA arranged in a Cartesian Grid and as such each ADPLL will be referred to using matrix indexing corresponding to their grid location. For each configuration three different modes of operation were examined: Each ADPLL locking to the same 5 MHz external reference, each ADPLL connected to the left and above neighbours in a ``uni-directional'' network and finally a true ADPLL network where all neighbours are connected. For comparing each configuration gains which had previously been determined to provide the best jitter performance were chosen. The data was captured using an Agilent MSO7054 A at 4 GSa/s. Each point of data is the mean of five 200 microsecond long captures which were then analysed in Matlab, and Table \ref{table:b_c_perf} displays the results of this analysis.\\
The first columns of the table relate to the cycle-to-cycle jitter of the system, the difference between sequential distances, and the second portion gives the Time Interval Error, or difference between the actual timing of the edge and the ``ideal'' time as set by the reference. This combination of measurements will show how well each design performs the task of staying consistent from cycle to cycle both in relation to itself and the system at large.

\begin{table}[!ht]
	\begin{center}
		\begin{tiny}
			\begin{tabular}{cc|c|c|c|c|c|c|}           
				\cline{3-8}
				&& \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Time Interval Error. (ns)} \T\\
				\cline{3-8}	
				&&PLL 11&PLL 12&PLL 22&PLL 11&PLL 12&PLL 22\T\\
				\hline
				\multicolumn{2}{|l|}{Osc 1, PDET A}&-&-&-&-&-&-\T\\
				\multicolumn{2}{|r|}{Free PLLs}&1.5988&1.6178&1.6220&&& \T\\
				\multicolumn{2}{|r|}{Uni-dir.}&1.9804&2.0316&2.1317&&& \T\\
				\multicolumn{2}{|r|}{Bi-dir.}&2.0124&2.0681&2.0732&&& \T\\
				\hline
				\multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
				\multicolumn{2}{|r|}{Free PLLs}&0.62499 $\mu$s&0.75919 $\mu$s&0.79049  $\mu$s&&& \T\\
				\multicolumn{2}{|r|}{Uni-dir.}&0.76822 $\mu$s&0.9009 $\mu$s&1.1407 $\mu$s&&& \T\\
				\multicolumn{2}{|r|}{Bi-dir.}&0.70393 $\mu$s&0.70201 $\mu$s&1.241 $\mu$s&&& \T\\
				\hline
				\multicolumn{2}{|l|}{Osc 2, PDET B}&-&-&-&-&-&-\T\\
				\multicolumn{2}{|r|}{Free PLLs}&4.1680&3.9803&5.0741&&& \T\\
				\multicolumn{2}{|r|}{Uni-dir.}&4.3820&5.7315&6.2336 &&& \T\\
				\multicolumn{2}{|r|}{Bi-dir.}&5.6137&7.1831&7.1612 &&& \T\\
				\hline
				\B
			\end{tabular}
		\end{tiny}
		\caption{}
		\label{table:b_c_perf}
	\end{center}
\end{table}

From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation. In this system the FPGA clock is at 258 MHz, the maximum possible, thus giving a detection resolution of 4 nanoseconds. This is 4 times greater than the period resolution of approximately 1.2 nano seconds of the oscillator, and requires the gains of the PI filter to be increased fourfold in order to respond appropriately to the error detected. The problems of relatively inaccurate detection and increased gains combine to produce a system that is only able to achieve poor locking performance.
Both other combinations work well however, delivering a level of jitter that is both low yet allows an indication how impactful variations in architecture or control parameters to be seen visually on an oscilloscope.\\
\begin{figure}[h]
	\centering
	\includegraphics[width=0.25\textwidth]{distrib_ring}
	\caption{Distribution of Periods for Inverter Based Design.}
	\label{fig:dist1}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.25\textwidth]{distrib_pa}
	\caption{Distribution of Periods for FPGA Clocked Design.}
	\label{fig:dist2}
\end{figure}
Figures \ref{fig:dist1} \& \ref{fig:dist2} paint a picture of the unrealistic nature of the FPGA clocked designs as in comparison to the ring based design each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment. This effects the suitability of the FPGA clocked designs for the improvement of simulation models or the examination of how a novel block maybe impact the behaviour of the network.
\section*{Conclusion}
An FPGA is a suitable environment for the prototyping of ADPLL networks and can fill a number of roles depending on the designer's requirements. However the designer should bear in mind the platform's limitations especially when attempting to use a frequency that is not orders of magnitude lower than the maximum clock available from the FPGA's clock manager. Inverter based and FPGA clocked designs both have a niche in which they are the preferred option, with the clocked design better suiting replication of an existing design at scaled down frequencies and the inverter based designs better in the modelling and analysis of ADPLL network behaviour.

\bibliographystyle{ieeetran}
\bibliography{conf}
